A New Dual Terminal Sliding Mode Control Design For Rigid Robotic Manipulator

### 2016-05-29 paper3_manuscript.tex
20160529

This paper proposes a novel dual terminal sliding mode control for rigid robotic manipulator. A new terminal sliding mode manifold design technique is addressed to achieve finite time convergence of the sliding phase, and furthermore utilizing the involved dual variables combines the individual sliding mode surfaces to obtain a integral sliding mode dynamics, the convergent time of which is also integral for the entire system. The underactuated issue is coped with by introducing the hierarchical sliding mode methodology into the basis dual sliding mode design. These newly proposed methods also have the characteristics of nonsingularity and chattering suppression, and the effectiveness and high efficiency are verified by stabilizing the motions of the overhead crane and the rigid robotic manipulator. Simulation results validate the theoretical analysis about the proposed method.

@@

Variable structure systems (VSS) are best known due to their insensitivity to parameter variations and uncertainties~\cite{slotine1991applied}. Its widely utilization in practice case, such as DC and AC motors, process control, aircrafts, satellites, robots and so on, has validated robustness of VSS. As one of the most important branch of the VSS, the sliding mode control (SMC) has become a popular solution to the issue about various control and signal process missions in recent years~\cite{zhao2015nonlinear,zhang2015attitude}. Meza-S{\'a}nchez et al. applied the second order sliding mode controller to a 3-DOF helicopter prototype, and verified the effectiveness via the experimental test bed~\cite{meza2015output}. Zheng et al. implemented the SMC to eliminate the effect of the input quantization generated between the coder and the decoder~\cite{zheng2016sliding}. Guzman et al. investigated the application of a SMC in a power factor rectifier with constant switching frequency~\cite{guzman2016sliding}. Said et al. designed a robust adaptive sliding mode method to guarantee the asymptotically stability of a boost converter~\cite{oucheriah2013pwm}. Roughly speaking, the SMC scheme is competent for normal scientific mission, due to its insensitivity of parameters and uncertainties in system, however, there still exist some obstacles, such as noncontinuous switching input in high frequency and convergent speed design~\cite{boiko2013chattering,lee2009chattering}, preventing it from being utilized in some practical engineering cases~\cite{fridman2011sliding}. \par
For alleviating the side effects, many researchers addressed remarkable approaches to improve the performance of the SMC scheme in last decade~\cite{zong2013quasi,santiesteban2013time,mu2015continuous,evangelista2013lyapunov,gonzalez2014chattering,dadras2012fractional,zhao2013output}. Among these methods, the high order sliding mode control and super-twisting sliding mode control efforts to generate the continuous inputs~\cite{castillo2015higher,palosz2015laser,edwards2016adaptive,zhao2015finite,liu2015second}, which eliminate the chattering by hiding the switching in high-order terms or absorbing the switching via some techniques. \par
Besides the issue on repressing the chattering, the fast convergence of the SMC scheme is also a hot subject. Normal linear sliding mode control takes two steps to solve a problem about the stability, namely, conducting the system states onto the prescript sliding manifold and driving these states to the origin along the manifold, and some accelerating operations usually occur in this two phases. The first step is said to be the approaching phase which usually requires the states converge to the desired manifold in finite time, and selecting suitable Lyapunov function and control parameters may accelerate this process; the second step is the sliding phase which is related to the designed manifold closely, and normal manifold design techniques are able to guarantee the asymptotically convergence. For achieving a fast convergence in the sliding phase, the terminal sliding mode (TSM) control is developed based on the finite time control theory~\cite{haimo1986finite,bhat1997finite}, and comparing with the sliding mode control design mentioned above, the most significant contributions of the terminal sliding mode is introducing the finite time convergence to the sliding phase to complete the global finite time convergence. Owing to the global finite time convergence, the terminal sliding mode control has been utilized to achieve the high precise control, i.e., in~\cite{li2015robust}, an adaptive TSM scheme controlled a $2$-link robotic manipulator to track the variable signal. Feng et al. presented a global terminal sliding mode algorithm for manipulating the rigid robot, and produced the control design by using Lagrangian dynamic expression~\cite{feng2002non}. In~\cite{kamal2016continuous}, Kamal addressed a continuous homogeneous sliding mode algorithm for a perturbed second-order plant, and the algorithm could be regarded as a combination of terminal and super-twisting sliding-mode, providing a convergence control in finite time. \par

@@

In this paper, a dual sliding mode (DTSM) controller is proposed for a class of second-order nonlinear systems with disturbances and uncertainties. A novel DTSM manifold is exploited to fulfill the finite time convergence, and the approaching and sliding time are both able to be calculated directly. For coping with the underacutated issue, cooperating with the hierarchical sliding surface design methodology, we develop a hierarchical dual terminal sliding mode (HDTSM) control. A modified DTSM (MDTSM) control scheme is presented to stabilize the motion of the $n$-link robotic manipulators. The remained paper is organized as follows: Section~\ref{sec:2} introduces a DTSM scheme for a class of second-order system, and the finite time convergence of the approaching and sliding is verified. In addition, we add performance parameters for improving the DTSM scheme to obtain the MDTSM scheme, and the newly presented method has a better performance in theoretical analysis; for stabilizing the underactuated system, the HDTSM design is developed as well in this section. In Section~\ref{sec:3}, the MDTSM is applied to the motion regulation of the $n$-link robotic  manipulator, and the stability analysis is also carried out. Section~\ref{sec:4} exhibited the simulation results, and finally the conclusion is assigned in Section~\ref{sec:5}.

@@

where $\bm x = [x_1,x_2,x_3,x_4]^T$ is the system vector state; $f(\bm x,t)$ and $b(\bm x,t)\neq 0$ are smooth nonlinear functions with respect to $\bm x$; $u_i(t)$, $i=1,2$ stands for the scalar input; $g_i(\bm x,t)$, $i=1,2$  is defined as a bounded nonlinear function which satisfies $\Vert g_i(\bm x,t)\Vert\le l_i$, $l_i>0$, meaning the uncertainty and disturbance of the system. For the system mentioned above, we can establish a desired dual terminal sliding manifold as:

@@

where $sgn(x)$ is the sign function, and $0<\alpha<1$ is a positive constant. For the sake of simplicity, the symbol $t$ is usually omitting in this paper, which does not cause any confusion, and it will appear once the derivation concerns for the factor of the time strictly, i.e., the desired sliding manifold $\bm s(t)$ is often abbreviated to $\bm s$.\par

@@

On the basis of Theorem \ref{theorem:1}, we proceed to develop the parameters in the manifold to improve the performance, which introduces a modified method for stabilizing the system~(\ref{eq:second-order system}), and furthermore this method makes the design of the manifold more flexible. Consider the system~(\ref{eq:second-order system}) with the manifold manifold $\bm s = [s_1,s_2]^T$, the details of which are described by following expressions:

@@

\section{Dual terminal sliding mode control for rigid manipulators}\label{sec:3}
In this section, a DTSM control is proposed for the rigid $2$-link robotic manipulator

@@

\bm \varepsilon = \bm q -\bm q_d

@@

At this point, the DTSM scheme for $2$-link robotic manipulator has been demonstrated, and this technique is able to be extended to $n$-link manipulator. Consider a $n$-link robotic manipulator in which $n$ is an even number, and the desired sliding manifold follows

@@

Referring to Theorem \ref{theorem:4}, it's easy to obtain the DTSM input for the even $n$-link manipulator. Essentially, this technique requires that the sliding mode manifold has to consist of $n/2$ pairs of individual sliding surfaces, in another words, this design for the manifold can be regarded as a work for yielding manifolds of the $n/2$ sub-systems.\par
Considering the situation that $n$ is an odd number, the $n$-link manipulator as we know is assumed as a full-actuated system, and hence an individual dimension is able to be excluded by using conventional terminal sliding mode control, which results in that the remained $n-1$-link manipulator can be controlled by the proposed DTSM scheme.

@@

The singularity of the DTSM scheme for the robotic manipulator can be removed by using the technique introduced in Theorem \ref{theorem:3}.

@@

and a MDTSM control scheme is produced conveniently. It's obvious that the disturbance in the system~(\ref{eq:simulation second-order system}) behaves bounded, which is said to be the following uncertainty terms:

@@

From Figure~\ref{Figure:2}, it's apparent that the states converge to their desired states. Figure~\ref{Figure:3} demonstrates the inputs, the amplitudes of which are closely related to $\eta$, for the purpose of avoiding the largely chattering caused by $\eta$, the saturation function $sat()$ is applied to take place of $sign()$ to absorb it. Also, although the states have be adjusted to the origin and stay there in subsequent time, one can find that the inputs don't maintain constant but fluctuate in a certain range. The nature of the situation mentioned above is the input compensating the error caused by the distance.\par
At this point, some effectiveness of the MDTSM control scheme proposed above has been verified, and in the next section, we will apply it to stabilize the behavior of a $2$-link robotic manipulator directly.

@@

\begin{figure}
\centering
\includegraphics[width=0.7\textwidth]{paper3_fig3.eps}
\caption{System inputs}
\label{Figure:3}
\end{figure}
\subsection{Uneractuated example}
Figure~\ref{Figure:8} shows an overhead crane system, the state space equations of which is same with the formula in Eq.~(\ref{eq:underactuated system}) with state variables selected as $\bm x=(x,\dot x,q,\dot q)$, and the details are listed:
\begin{align}
\begin{split}
f_1&=(M+m\sin^2q)^{-1}(mL\dot q^2\sin q+mg\sin q\cos q)\\
b_1&=(M+m\sin^2q)^{-1}\\
f_2&=-(ML+mL\sin^2q)^{-1}((M+m)g\sin q+mL\dot q^2\sin q\cos q)\\
b_2&=-(ML+mL\sin^2q)^{-1}\cos q,
\end{split}
\end{align}
where $x$ is defined as the displacement of the horizontal direction, and $q$ stands for the swing angle from the vertical line. $M=1kg$ and $m=0.8kg$ donate the masses of the trolley and the load in Figure~\ref{Figure:8}, respectively, and $g=9.8m/s^2$ refers to as the acceleration of gravity. $L=0.305m$ represents the length of the rope. In this simulation, the disturbances are also introduced:
\begin{align} 
g_1(\bm x)&= 0.02\sin t\\
g_2(\bm x)&=0.07e^{-4t},
\end{align}
and those boundaries are easy to obtain. The initial condition of the overhead crane system is $\bm x_0 = (0.1,0,0.1,0)$, while the desired status is $\bm x_d = (2,0,0,0)$. The HDSM controller can be designed with the following set of parameters:
$c_1=1$, $c_2=4$, $\eta_u = 0.1$, $k_u=1$, $\gamma = 1$, $\lambda = 3$, $\alpha = 0.6$.\par
Figure~\ref{Figure:8} shows the state response curves of the overhead crane system controlled using the HDSM control scheme, and the results show that the controller stabilize the displacement at $6s$ approximately, though the disturbance exists in the system. One can find that the swing angle is fluctuating in a fairly small range, the boundary of which is not more than $1\times 10^-3rad$, hence this phenomena is reasonable and also occurs in the stable hierarchical sliding mode control \cite{wang2004design}.  
\begin{figure}
\centering
\includegraphics[width=0.3\textwidth]{paper3_fig8.eps}
\caption{Overhead crane schematic}
\label{Figure:8}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.7\textwidth]{paper3_fig9.eps}
\caption{Overhead crane schematic}
\label{Figure:9}
\end{figure}
\subsection{Performance of the robotic manipulator}
In order to evaluate the performance of the proposed control scheme, a simulation about the $2$-link robotic manipulator is presented in the following context. Figure~\ref{Figure:1} is the $2$-link manipulator schematic, the dynamics of which can be explained by the dynamic equations in following forms:
\begin{figure}
\centering
\includegraphics[width=0.3\textwidth]{paper3_fig1.eps}
\caption{$2$-link manipulator schematic}
\label{Figure:1}
\end{figure}

>>>>

This paper proposes a novel dual terminal sliding mode control for rigid robotic manipulator. A new terminal sliding mode manifold design technique is addressed to achieve finite time convergence of the sliding phase, and furthermore utilizing the involved dual variables combines the individual sliding mode surfaces to obtain an integral sliding mode dynamics, the convergent time of which is also integral for the entire system. The underactuated issue is coped with by introducing the hierarchical sliding mode methodology into the basic dual sliding mode design. These newly proposed methods also have the characteristics of nonsingularity and chattering suppression, and the effectiveness and high efficiency are verified by stabilizing the motions of the overhead crane and the rigid robotic manipulator. Simulation results validate the theoretical analysis about the proposed method.

@@

Variable structure systems (VSS) are best known due to their insensitivity to parameter variations and uncertainties~\cite{slotine1991applied}. Its widely utilization in reality, such as DC and AC motors, process control, aircrafts, satellites, robots and so on, has validated robustness of VSS. As one of the most important branch of the VSS, the sliding mode control (SMC) has become a popular solution to the issue about various control and signal process missions in recent years~\cite{zhao2015nonlinear,zhang2015attitude}. Meza-S{\'a}nchez et al. applied the second-order sliding mode controller to a 3-DOF helicopter prototype, and verified the effectiveness via the experimental test bed~\cite{meza2015output}. Zheng et al. implemented the SMC to eliminate the effect of the input quantization generated between the coder and the decoder~\cite{zheng2016sliding}. Guzman et al. investigated the application of a SMC in a power factor rectifier with constant switching frequency~\cite{guzman2016sliding}. Said et al. designed a robust adaptive sliding mode method to guarantee the asymptotically stability of a boost converter~\cite{oucheriah2013pwm}. Roughly speaking, the SMC scheme is competent for normal scientific mission, due to its insensitivity of parameters and uncertainties in system, however, there still exist some obstacles, such as noncontinuous switching input in high frequency and convergent speed design~\cite{boiko2013chattering,lee2009chattering}, preventing it from being utilized in some practical engineering cases~\cite{fridman2011sliding}. \par
For alleviating the side effects, many researchers addressed remarkable approaches to improve the performance of the SMC scheme in last decade~\cite{zong2013quasi,santiesteban2013time,mu2015continuous,evangelista2013lyapunov,gonzalez2014chattering,dadras2012fractional,zhao2013output}. Among these methods, the high-order sliding mode control and super-twisting sliding mode control effort to generate the continuous inputs~\cite{castillo2015higher,palosz2015laser,edwards2016adaptive,zhao2015finite,liu2015second}, which eliminate the chattering by hiding the switching in high-order terms or absorbing the switching via some techniques. \par
Besides the issue on repressing the chattering, the fast convergence of the SMC scheme is also a hot subject. Normal linear sliding mode control takes two steps to solve a problem about the stability, namely, conducting the system states onto the prescript sliding manifold and driving these states to the origin along the manifold, and some accelerating operations usually occur in this two phases. The first step is said to be the approaching phase which usually requires the states converge to the desired manifold in finite time, and selecting suitable Lyapunov function and control parameters may accelerate this process; the second step is the sliding phase which is related to the designed manifold closely, and trivial manifold design techniques are able to guarantee the asymptotically convergence. For achieving a fast convergence in the sliding phase, the terminal sliding mode (TSM) control is developed based on the finite time control theory~\cite{haimo1986finite,bhat1997finite}, and comparing with the sliding mode control design mentioned above, the most significant contributions of the terminal sliding mode is introducing the finite time convergence to the sliding phase to complete the global finite time convergence. Owing to the global finite time convergence, the terminal sliding mode control has been utilized to achieve the high precise control, i.e., in~\cite{li2015robust}, an adaptive TSM scheme controlled a $2$-link robotic manipulator to track the variable signal. Feng et al. presented a global terminal sliding mode algorithm for manipulating the rigid robot, and produced the control design by using Lagrangian dynamic expression~\cite{feng2002non}. In~\cite{kamal2016continuous}, Kamal addressed a continuous homogeneous sliding mode algorithm for a perturbed second-order plant, and the algorithm could be regarded as a combination of terminal and super-twisting sliding-mode, providing a convergence control in finite time. \par

@@

In this paper, a dual terminal sliding mode (DTSM) controller is proposed for a class of second-order nonlinear systems with disturbances and uncertainties. A novel DTSM manifold is exploited to fulfill the finite time convergence, and the approaching and sliding time are both able to be calculated directly. For coping with the underacutated issue, cooperating with the hierarchical sliding surface design methodology, we develop a hierarchical dual terminal sliding mode (HDTSM) control. A modified DTSM (MDTSM) control scheme is presented to stabilize the motion of the $n$-link robotic manipulators. The remained paper is organized as follows: Section~\ref{sec:2} introduces a DTSM scheme for a class of second-order system, and the finite time convergence of the approaching and sliding phase is verified. In addition, we add performance parameters for improving the DTSM scheme to obtain the MDTSM scheme, and the newly presented method has a better performance in theoretical analysis; for stabilizing the underactuated system, the HDTSM design is developed as well in this section. In Section~\ref{sec:3}, the MDTSM is applied to the motion regulation of the $n$-link robotic  manipulator, and the stability analysis is also carried out. Section~\ref{sec:4} exhibits the simulation results, and finally the conclusion is assigned in Section~\ref{sec:5}.

@@

where $\bm x = [x_1,x_2,x_3,x_4]^T$ is the system vector state; $f(\bm x,t)$ and $b(\bm x,t)\neq 0$ are smooth nonlinear functions with respect to $\bm x$; $u_i(t)$, $i=1,2$ stands for the scalar input; $g_i(\bm x,t)$, $i=1,2$  is defined as a bounded nonlinear function which satisfies $\Vert g_i(\bm x,t)\Vert\le l_i$, $l_i>0, i=1,2$, meaning the uncertainty and disturbance of the system. For the system mentioned above, we can establish a desired dual terminal sliding manifold as:

@@

where $sgn(x)$ is the sign function, and $0<\alpha<1$ is a positive constant. For the sake of simplicity, the symbol $t$ is usually omitted in this paper, which does not cause any confusion, and it will appear once the derivation concerns for the factor of the time strictly, i.e., the desired sliding manifold $\bm s(t)$ is often abbreviated to $\bm s$.\par

@@

On the basis of Theorem \ref{theorem:1}, we proceed to develop the parameters in the manifold to improve the performance, which introduces a modified method for stabilizing the system~(\ref{eq:second-order system}), and furthermore this method makes the design of the manifold more flexible. Consider the system~(\ref{eq:second-order system}) with the manifold $\bm s = [s_1,s_2]^T$, the details of which are described by following expressions:

@@

\section{Modified Dual terminal sliding mode control for rigid manipulators}\label{sec:3}
In this section, the MDTSM control is proposed for the rigid $2$-link robotic manipulator

@@

\bm \varepsilon = \bm q -\bm q_d.

@@

At this point, the MDTSM scheme for $2$-link robotic manipulator has been demonstrated, and this technique is able to be extended to $n$-link manipulator. Consider a $n$-link robotic manipulator in which $n$ is an even number, and the desired sliding manifold follows

@@

Referring to Theorem \ref{theorem:4}, it's easy to obtain the MDTSM input for the even $n$-link manipulator. Essentially, this technique requires that the sliding mode manifold has to consist of $n/2$ pairs of individual sliding surfaces, in another words, this design for the manifold can be regarded as a work for yielding manifolds of the $n/2$ sub-systems.\par
Considering the situation that $n$ is an odd number, the $n$-link manipulator as we know is assumed as a full-actuated system, and hence an individual dimension is able to be excluded by using conventional terminal sliding mode control, which results in that the remained $n-1$-link manipulator can be controlled by the proposed MDTSM scheme.

@@

The singularity of the MDTSM scheme for the robotic manipulator can be removed by using the technique introduced in Theorem \ref{theorem:3}.

@@

and an MDTSM control scheme is produced conveniently. It's obvious that the disturbance in the system~(\ref{eq:simulation second-order system}) behaves bounded, which is said to be the following uncertainty terms:

@@

From Figure~\ref{Figure:2}, it's apparent that the states converge to their desired states. Figure~\ref{Figure:3} demonstrates the inputs, the amplitudes of which are closely related to $\eta$, for the purpose of avoiding the largely chattering caused by $\eta$, the saturation function $sat()$ is applied to take place of $sign()$ to absorb it. Also, although the states have been adjusted to the origin and stay there in subsequent time, one can find that the inputs don't maintain constant but fluctuate in a certain range. The nature of the situation mentioned above is the input compensating the error caused by the distance.\par
At this point, some effectiveness of the MDTSM control scheme proposed above has been verified, and in Section \ref{sec:robotica manipulator}, we will apply it to stabilize the behavior of a $2$-link robotic manipulator directly.

@@

\begin{figure}
\centering
\includegraphics[width=0.7\textwidth]{paper3_fig3.eps}
\caption{System inputs}
\label{Figure:3}
\end{figure}
\subsection{Uneractuated example}
Figure~\ref{Figure:8} shows an overhead crane system, the state space equations of which is same with the formula in Eq.~(\ref{eq:underactuated system}) with state variables selected as $\bm x=(x,\dot x,q,\dot q)$, and the details are listed:
\begin{align}
\begin{split}
f_1&=(M+m\sin^2q)^{-1}(mL\dot q^2\sin q+mg\sin q\cos q)\\
b_1&=(M+m\sin^2q)^{-1}\\
f_2&=-(ML+mL\sin^2q)^{-1}((M+m)g\sin q+mL\dot q^2\sin q\cos q)\\
b_2&=-(ML+mL\sin^2q)^{-1}\cos q,
\end{split}
\end{align}
where $x$ is defined as the displacement of the horizontal direction, and $q$ stands for the swing angle from the vertical line. $M=1kg$ and $m=0.8kg$ donate the masses of the trolley and the load in Figure~\ref{Figure:8}, respectively, and $g=9.8m/s^2$ refers to as the acceleration of gravity. $L=0.305m$ represents the length of the rope. In this simulation, the disturbances are also introduced:
\begin{align}
g_1(\bm x)&= 0.02\sin t\\
g_2(\bm x)&=0.07e^{-4t},
\end{align}
and those boundaries are easy to obtain. The initial condition of the overhead crane system is $\bm x_0 = (0.1,0,0.1,0)$, while the desired status is $\bm x_d = (2,0,0,0)$. The HDSM controller can be designed with the following set of parameters:
$c_1=1$, $c_2=4$, $\eta_u = 0.1$, $k_u=1$, $\gamma = 1$, $\lambda = 3$, $\alpha = 0.6$.\par
Figure~\ref{Figure:8} shows the state response curves of the overhead crane system controlled using the HDSM control scheme, and the results show that the controller stabilize the displacement at $6s$ approximately, though the disturbance exists in the system. One can find that the swing angle is fluctuating in a fairly small range, the boundary of which is not more than $1\times 10^-3rad$, hence this phenomena is reasonable and also occurs in the stable hierarchical sliding mode control \cite{wang2004design}.
\begin{figure}
\centering
\includegraphics[width=0.3\textwidth]{paper3_fig8.eps}
\caption{Overhead crane schematic}
\label{Figure:8}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.7\textwidth]{paper3_fig9.eps}
\caption{Overhead crane schematic}
\label{Figure:9}
\end{figure}
\subsection{Performance of the robotic manipulator}\label{sec:robotica manipulator}
In order to evaluate the performance of the proposed control scheme, a simulation about the $2$-link robotic manipulator is presented in the following context. Figure~\ref{Figure:1} is the $2$-link manipulator schematic, the dynamics of which can be explained by the dynamic equations in following forms:
\begin{figure}
\centering
\includegraphics[width=0.3\textwidth]{paper3_fig1.eps}
\caption{$2$-link manipulator schematic}
\label{Figure:1}
\end{figure}
